---
title: "team"
author: "Danny Nguyen and Connor Hergenreter"
date: "March 26, 2019"
output: html_document
---

```{r setup}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readxl)
defense <- read_excel("cyclonesFootball2018.xlsx", sheet='Defensive')
offense <- read_excel("cyclonesFootball2018.xlsx", sheet='Offensive')
biography <- read_excel("cyclonesFootball2018.xlsx", sheet='Biography')
```

```{r}
#Part One: Cleaning data 
#Question 1: The player names and opponent names are factors.
defense <- defense %>% 
  mutate(Name = as.factor(Name),
         Opponent_Opponent = as.factor(Opponent_Opponent)) 
str(defense)

offense <- offense %>%
  mutate(Name = as.factor(Name),
         Opponent_Opponent = as.factor(Opponent_Opponent)) 
str(offense)

biography <- biography %>% 
  mutate(Name = as.factor(Name))
str(biography)
```

```{r}
#Question 2: All offensive and defensive statistics are numerical. The Weight column in biogrpahy is also numeric.
defense <- mutate_at(defense, vars(Tackles_Solo, Tackles_ASST, Tackles_TFL, Tackles_Sack, Turnover_FF, Turnover_FR, Turnover_INT, Pass_QBH, Pass_PB), as.numeric)
str(defense)

offense <- mutate_at(offense, vars(Receiving_REC, Receiving_YDS, Receiving_TD, Rushing_ATT, Rushing_YDS, Rushing_TD, `Passing_CMP-ATT`, Passing_YDS, Passing_TD, Passing_INT), as.numeric)
str(offense)

biography <- biography %>% 
  mutate(Weight = as.numeric(Weight))
str(biography)
```

```{r}
#Question 3: Bonus: Change the Height column in biogrpahy into numeric.
biography <- biography %>%
  mutate(Height = as.numeric(Height), na.rm = TRUE)
str(biography)

#Create and print defClean, offClean, and bioClean
defClean <- defense
offClean <- offense
bioClean <- biography

print(defClean)
print(offClean)
print(bioClean)
```

#Part 2

#Problem 1-2

Everything looks rare relative to Tackles_ASST and Tackles_Solo. Pass_QBH, Pass_PF, Tackles_Sack, and Tackles_TFL max out at three and all look pretty rare. The rarest three are the Turnover_FF, Turnover_FR, and Turnover_INT. Those three only have players getting one per game if any.

```{r}
defTid <- defClean %>% gather(key=Statistic, value=Count, 3:11)

ggplot(defTid, aes(x=Count)) + geom_histogram(binwidth=1) + facet_wrap(~Statistic) + scale_x_continuous(breaks=seq(0,10,1)) + labs(title="Defensive Statistic Counts", x="Player Max", y="Count")
```

#Problem 3

Based off this scatterplot it appears that the Iowa State players did better tackle-wise against West Virgina. Too many players got no tackles against Kansas State for this to be much of a competition

```{r}
wvks <- defClean %>% filter(Opponent_Opponent=="West Virginia" | Opponent_Opponent=="Kansas State") %>% select(Name, Opponent_Opponent, Tackles_Solo) %>% spread(Opponent_Opponent, Tackles_Solo) %>% mutate_all(funs(replace(., is.na(.), 0)))

print(tbl_df(wvks), n=500)

ggplot(wvks, aes(x=`West Virginia`, y=`Kansas State`)) + geom_point() + scale_y_continuous(breaks=seq(0,9,1)) + labs(title="Player Tackle Counts Against Kansas State and West Virginia", x="West Virginia", y="Kansas State")
```

#Part 2 of Part 2

```{r}
bioSep <- bioClean %>% separate(Hometown, c("Town", "State"), sep=",") %>% group_by(State) %>% summarize(n())
```

#Part 3

```{r}
offBio <- inner_join(offClean, bioClean, by="Name")
print(tbl_df(offBio), n=5)
```

There does seem to be a bit of a relationship with weight and receiving yards. Lighter players get a decent amount of receiving yards but when weight hits 225 it really skyrockets. If we exclude the 225 weight group then there is a negative relationship between weight and receiving yards.

```{r}
ggplot(offBio, aes(x=Weight, y=Receiving_YDS)) + geom_point() + labs(Title="Receiving Yards by Weight", x="Weight", y="Receiving Yards")
```

